A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-18 graph_syntax.tex
stuff

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. So we define an expression block, as the collection of variable and model 
nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

@@

  Note that the variable $x_1$ from the Sellar problem is usually refered to as 
  a local variable, which is contradicted by the given graph. This discrepancy 
  arises from the explicit treatment of objectives and constraints as separate 
  expression blocks. Normally in the Sellar problem $x_1$ is considered a local 
  variable because it only directly affects analysis block $D_1$. By forcing
  the expansion of $F$ into an expression block with its own inputs, then $F$ 
  gets it's own copy of the $x_1$ variable. This necessitates the creation 
  of the global $x_1^*$ node in the graph with two outgoing edges to link the 
  two $x_1$ inputs nodes in the different blocks. Interestingly, using this 
  definition, where $x_1^*$ is a global variable fits nicely within 
  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun_thesis}. 
  The rules of setting up a problem with CO require that each discipline be 
  given an independent local copy of all global variables, while any local 
  variables are retained uniquely within their respective disciplines. But 
  when a local design variable shows up explicitely in the objective function 
  or global constraint then a global variable is created, and the discipline again 
  gets an independent local copy of it. Given our definition of global vs local
  variable nodes, no such exception for local variables is necessary. When you 
  include a given variable in an expression node, it becomes global by definition 
  and should be treated as such in CO.

@@

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{XDSM/simple}
      [[insert graph here]]
      \caption{A simple graph with local and global nodes \label{f:local_global}}
      \end{center}
  \end{figure}

  Simply stating that any node or block with incomming or outgoing edges
  from more than one other node or block is global works glosses over a more 
  subtle aspect of some MDAO problems. The locality of any node in a graph is 
  a relative property. For instance, a single node might have outgoing edges 
  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. Then 
  you could say that this node was global with respect to $A$ and $B$, but local 
  to the group $A,B$ with respect to $C$. This situation produces a natural 
  hierarchy in a graph, where you could form increasingly smaller groups as you 
  segment problem into finer localities.

  When solving a problem using a heirarchical decomposition approach, 
  there are various techniques for partitioning that are effective in different 
  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,
  sobieszczanski1997,Perez2004,allison2009optimal}. Tosserams et al proposed a 
  language based syntax for describing the partitioning of problems, named $\Psi$
  \cite{tosserams2010specification}. $\Psi$ provides an opposing perspective to 
  this work, where they compose hierarchies from the bottom up, into larger and 
  larger groups. However, $\Psi$ provides a compiler that can produce a standard 
  problem representation of the assembled problem, and a number of example converters
  from that standard representation to application specific formats. It would be 
  possible to provide such a converter to build a graph form the $\Psi$ format.

>>>>

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. So we define an expression block, as the collection of variable and model 
nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
highlights the objective and constraint blocks from the Sellar problem.

@@

  Note that the variable $x_1$ from the Sellar problem is usually refered to as 
  a local variable, which is contradicted by the given graph. This discrepancy 
  arises from the explicit treatment of objectives and constraints as separate 
  expression blocks. Normally in the Sellar problem $x_1$ is considered a local 
  variable because it only directly affects analysis block $D_1$. By forcing
  the expansion of $F$ into an expression block with its own inputs, then $F$ 
  gets it's own copy of the $x_1$ variable. This necessitates the creation 
  of the global $x_1^*$ node in the graph with two outgoing edges to link the 
  two $x_1$ inputs nodes in the different blocks. Interestingly, using this 
  definition, where $x_1^*$ is a global variable fits nicely within 
  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun1996thesis}. 
  The rules of setting up a problem with CO require that each discipline be 
  given an independent local copy of all global variables. All local 
  variables are retained uniquely within their respective disciplines, except 
  when a local design variable shows up explicitely in the objective 
  or global constraint. In that case a global variable is created, and the discipline again 
  gets an independent local copy of it. Given our definition of global vs local
  variable nodes, no such exception for local variables in global expressions is 
  necessary. When you include a given variable in an expression block, it becomes 
  global by definition and should be treated as such in CO.

@@

  Simply stating that any node or block with incomming or outgoing edges
  from more than one other node or block is global works glosses over a more 
  subtle aspect of some MDAO problems. The locality of any node in a graph is 
  a relative property. For instance, a single node might have outgoing edges 
  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. Then 
  you could say that this node was global with respect to $A$ and $B$ separately, but local 
  to the group $A,B$. This situation produces a natural 
  hierarchy in a graph, where you could form increasingly smaller groups as you 
  segment problem into finer localities.

  When solving a problem using a heirarchical decomposition approach, 
  there are various techniques for partitioning that are effective in different 
  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,
  sobieszczanski1997,Perez2004,allison2009optimal}. Tosserams et al proposed a 
  language based syntax for describing the partitioning of problems, named $\Psi$
  \cite{tosserams2010specification}. $\Psi$ provides an opposing perspective to 
  this work, where they compose hierarchies from the bottom up, into larger and 
  larger groups. However, $\Psi$ provides a compiler that can produce a standard 
  problem representation of the assembled problem, and a number of example converters
  from that standard representation to application specific formats.
